## **Regression and Model validation**


### **Data wrangling**



In this section, the data set given in this [link](http://www.helsinki.fi/~kvehkala/JYTmooc/JYTOPKYS3-data.txt) has been  preprocess for further/downstream analysis. After creating a folder named **'data'** in my IODS-project repository, using Rstudio a new R script named **'create_learning2014'** file created and write all the steps  used in the data wrangling process and saved in the **'data'** folder. All the steps I have done in the data wrangling preprocess can be find  [here](https://github.com/melakbet/IODS-project/blob/master/data/create_learning2014.R) and/or in the code shown below.


### **Analysis**

The data for this section is extracted from a survey conducted by **Kimmo Vehkalahti**  on teaching and learning approaches. The research project made possible by the Academy of Sciences funding to Teachers' Academy funding (2013-2015). The data was collected from 3.12.2014 - 10.1.2015.The surrvey includes 183 respondants and the questioner in toatal includes 60 variables/questions.

```{r, fig.width=11, fig.height=9, echo=FALSE}
#Melak Weldenegodguad
#date 15.11.2017
#install.packages("GGally")
#install.packages("GGally")
library(GGally)
library(ggplot2)
```

```{r}
students2014=read.table("data/learning2014.txt", sep="\t", header=TRUE)
dim(students2014)
str(students2014)

```

After preprocess the raw data,  the final data set have 7 column "gender, age, attitude, deep, stra, surf and points" and 166 indvidual as shown in the above run.

```{r, fig.width=11, fig.height=9}
plot_students2014=ggpairs(students2014, mapping = aes(col = gender,alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

plot_students2014

```

The above plot shown there are 110 femal and 56 male respondants in the survey. Moreover the plot shown how the variables (Points, Age, attitude, deep, stra and Surf) are correleted. Based on the absolute correlation value the attitude and points is higly correleted while deep learning and ponts are least corrleted.

The explanatory variable chosen  based on (absolute) correlation value and the top three explanatory variable chosen are attitude=  0.437, stra=  0.146 and surf= -0.144. Using this variables the model is build using the R function **"lm"**.

```{r, fig.width=11, fig.height=9}
my_model1 <- lm(Points ~ attitude + stra + surf, data = students2014)

summary(my_model1)

```


The significance paramter from the above summary table is the intercept  0.00322  and attitude (1.93e-08) ***. Hence stra nad surf are not condisdered in the model below. using again "**lm**" function linear regrression model build between points and attitude.The model after removing insignificant variables is summarized below. With regard to multiple r-squared value, we saw that value changed from 0.1927 (in older model) to 0.1856 (newer model). However,  F-Statistic (from 14.13 to 38.61) and p-value(3.156e-08 to 4.119e-09) have significantly improved. Thus, we can conclude that r-squared value may not always determine the quality of the model and the lower r-squared value might be due to outliers in the data. 

```{r}
my_model2 <- lm(Points ~ attitude , data = students2014)
summary(my_model2)
```

#### **Model Validation**

```{r, fig.width=11, fig.height=9}
par(mfrow = c(2,2))

plot(my_model2, which=c(1,2,5))

```


The three diagnostic model validation plots are shown above.The assumptions are

- The errors are normally distributed
- The errors are not correleted
- The errors have constant variance, 
- The size of a given error doesn't depend on the explanatory variables

Based on the above plots, we can conclude that the errors are normally distributed (clearly observed in q-q plot). Similarly, residual versus fitted model showed that the errors are not dependent on the attitude variable. Moreover, we can see that even two points (towards the right) have minor influence to the assumption in case of residual vs leverage model. All the models have adressed the outliers nicely. Thus, assumptions in all models are more or less valid.

